SUPPLEMENTARY INFORMATION: Overlay databank unlocks data-driven analyses of biomolecules for all

### 2023-12-19 SI.tex
Editorial updates

\renewcommand{\thetable}{S\arabic{table}}%
\renewcommand{\thefigure}{S\arabic{figure}}%
\renewcommand{\thesection}{S\arabic{section}}%
\renewcommand{\thepage}{S\arabic{page}}%

@@

\author[1]{Anne M. Kiirikki}         %ORCID: ?
\author[2,3]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
\author[2,4]{Lara S. Bort}             %ORCID: ?
\author[5]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
\author[6]{Fernando Favela-Rosales}
\author[7]{Tiago Mendes Ferreira}
\author[8,9]{Patrick F.J. Fuchs}
\author[10]{Rebeca Garcia-Fandino}
\author[11]{Ivan Gushchin}
\author[12,13]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
\author[14]{Norbert Ku{\v c}erka}
\author[15]{Patrik Kula}
\author[16]{Milla Kurki}
\author[11]{Alexander Kuzmin}
\author[26]{Anusha Lalitha} %ORCID: 0000-0002-2817-661X
\author[27,28]{Fabio Lolicato}
\author[17,18]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
\author[2,19,20]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
\author[25]{Cedric Mingham}
\author[29,30]{Luca Monticelli}
\author[1,31]{Ricky Nencini}
\author[19,20]{Alexey M. Nesterenko}
\author[21]{Thomas J. Piggot}
\author[22]{{\'A}ngel Pi{\~n}eiro}
\author[19,20]{Nathalie Reuter}
\author[12]{Suman Samantray} %ORCID: 0000-0003-3361-9582
\author[10,22,23]{Fabi{\'a}n Su{\'a}rez-Lest{\'o}n}
\author[19,20]{Reza Talandashti}
\author[1,24,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006

\affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
\affil[2]{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
\affil[3]{Department of Biomaterials, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
\affil[4]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
\affil[5]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, Jyv{\"a}skyl{\"a}, 40014, Finland}
\affil[6]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
\affil[7]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
\affil[8]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
\affil[9]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
\affil[10]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
\affil[11]{no affiliation}
\affil[12]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, 52428 Jülich, Germany}
\affil[13]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
\affil[14]{Department of Physical Chemistry of Drugs and Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}
\affil[15]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
\affil[16]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
\affil[17]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida 33612, United States of America}
\affil[18]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}
\affil[19]{Department of Chemistry, University of Bergen, 5020 Bergen, Norway}
\affil[20]{Computational Biology Unit, Department of Informatics, University of Bergen, 5020 Bergen, Norway}
\affil[21]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
\affil[22]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
\affil[23]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
\affil[24]{VTT Technical Research Centre of Finland, Espoo, Finland}
\affil[25]{Hochschule Mannheim, University of Applied Sciences, Mannheim, Germany}
\affil[26]{Institut Charles Gerhardt Montpellier (UMR CNRS 5253), Universit{\'e} Montpellier, Place Eug{\`e}ne Bataillon, 34095 Montpellier, Cedex 05, France}
\affil[27]{Heidelberg University Biochemistry Center, 69120 Heidelberg, Germany}
\affil[28]{Department of Physics, University of Helsinki, FI-00014 Helsinki, Finland}
\affil[29]{University of Lyon, CNRS, Molecular Microbiology and Structural Biochemistry (MMSB, UMR 5086), F-69007, Lyon, France}
\affil[30]{Institut National de la Sant{\'e} et de la Recherche M{\'e}dicale (INSERM), Lyon, France}
\affil[31]{Division of Pharmaceutical Biosciences, Faculty of Pharmacy, University of Helsinki, Helsinki 00014, Finland}

@@

\pagebreak
\section{Processing results in the application layer}

@@

\pagebreak
\section{Correlations of area per lipid and thickness with order parameters and form factors}

@@

\pagebreak
\section{Dependence of form factor and order parameters on the simulation box size}

@@

\pagebreak
\section{Finding the best models for PC and PE mixtures}

@@

\pagebreak
\section{Testing machine learning models against literature data}
\begin{figure}[hb]
    \centering
    \includegraphics[width=45mm]{Figures/MLtest/LinearRegressiontest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/Ridgetest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/MLPRegressortest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/Lassotest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/ElasticNettest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/DecisionTreeRegressortest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/RandomForestRegressortest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/GradientBoostingRegressortest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/AdaBoostRegressortest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/KNeighborsRegressortest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/XGBRegressortest.pdf}
    \caption{Predictions of area per lipids of multi-component membranes composed of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from linear regression model against literature data from simulations using CHARMM36 parameters (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}). Mean squared error (RMSE) and Pearson correlation coefficient are shown for each model. The black line depicts x=y where the prediction equals the literature value. Our machine learning models that are trained using simulations with wide range of force field parameters  have general tendency to predict larger area per lipid values than reported from CHARMM36 simulations in the literature. One explanation could be lower area per lipids predicted by the CHARMM36 parameters when compared with many other force fields in the training set~\cite{javanainen23}, yet other explanations cannot be ruled out.}
    \label{fig:MLTest}
\end{figure}

\pagebreak
\section{Water permeation through membranes}

@@

\pagebreak
\section{Water diffusion along membranes}

@@

%\newpage
\pagebreak
\section{Databank content}

@@

\clearpage
%\section{Experimental data}
\section{NMR experiments}
Acyl chain order parameters of POPE (Figs.~\ref{POPEexp1} and~\ref{POPEexp2}) and POPG (Figs.~\ref{POPGexp1} and~\ref{POPGexp2}) were analyzed from the same data that were previously recorded to determine headgroup order parameters~\cite{bacle21}. The analysis of the crowded spectral region at 29--31 ppm was based on the previous assignment reported for POPC membranes~\cite{ferreira13}. To measure the order parameters for DOPC (Fig.~\ref{fig:DOPCexp}), the sample was prepared and experiments performed similarly to previous studies~\cite{bacle21}.

>>>>

%\renewcommand{\thetable}{S\arabic{table}}%
%\renewcommand{\thefigure}{S\arabic{figure}}%
%\renewcommand{\thesection}{S\arabic{section}}%
%\renewcommand{\thepage}{S\arabic{page}}%

@@

\author[1]{Anne M. Kiirikki}         %ORCID: ?
\author[2,3]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
\author[2,4]{Lara S. Bort}             %ORCID: ?
\author[5]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
\author[6]{Fernando Favela-Rosales}
\author[7]{Tiago Mendes Ferreira}
\author[8,9]{Patrick F.J. Fuchs}
\author[10]{Rebeca Garcia-Fandino}
\author[11]{Ivan Gushchin}
\author[12,13]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
\author[14]{Norbert Ku{\v c}erka}
\author[15]{Patrik Kula}
\author[16]{Milla Kurki}
\author[11]{Alexander Kuzmin}
\author[17]{Anusha Lalitha} %ORCID: 0000-0002-2817-661X
\author[18,19]{Fabio Lolicato}
\author[20,21]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
\author[2,22,23]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
\author[24]{Cedric Mingham}
\author[25,26]{Luca Monticelli}
\author[1,27]{Ricky Nencini}
\author[22,23]{Alexey M. Nesterenko}
\author[28]{Thomas J. Piggot}
\author[29]{{\'A}ngel Pi{\~n}eiro}
\author[22,23]{Nathalie Reuter}
\author[12,30]{Suman Samantray} %ORCID: 0000-0003-3361-9582
\author[10,29,31]{Fabi{\'a}n Su{\'a}rez-Lest{\'o}n}
\author[22,23]{Reza Talandashti}
\author[1,32,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006

\affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
\affil[2]{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
\affil[3]{Department of Biomedicine, University of Bergen, 5020 Bergen, Norway}
\affil[4]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
\affil[5]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, Jyv{\"a}skyl{\"a}, 40014, Finland}
\affil[6]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
\affil[7]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
\affil[8]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), F-75005 Paris, France}
\affil[9]{Universit{\'e} Paris Cit{\'e}, F-75006 Paris, France}
\affil[10]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
\affil[11]{no affiliation}
\affil[12]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, 52428 Jülich, Germany}
\affil[13]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
\affil[14]{Department of Physical Chemistry of Drugs, Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}
\affil[15]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
\affil[16]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
\affil[17]{Institut Charles Gerhardt Montpellier (UMR CNRS 5253), Universit{\'e} Montpellier, Place Eug{\`e}ne Bataillon, 34095 Montpellier, Cedex 05, France}
\affil[18]{Heidelberg University Biochemistry Center, 69120 Heidelberg, Germany}
\affil[19]{Department of Physics, University of Helsinki, FI-00014 Helsinki, Finland}
\affil[20]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida 33612, United States of America}
\affil[21]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}
\affil[22]{Department of Chemistry, University of Bergen, 5007 Bergen, Norway}
\affil[23]{Computational Biology Unit, Department of Informatics, University of Bergen, 5008 Bergen, Norway}
\affil[24]{Hochschule Mannheim, University of Applied Sciences, Mannheim, Germany}
\affil[25]{University of Lyon, CNRS, Molecular Microbiology and Structural Biochemistry (MMSB, UMR 5086), F-69007, Lyon, France}
\affil[26]{Institut National de la Sant{\'e} et de la Recherche M{\'e}dicale (INSERM), Lyon, France}
\affil[27]{Division of Pharmaceutical Biosciences, Faculty of Pharmacy, University of Helsinki, Helsinki 00014, Finland}

\affil[28]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}

\affil[29]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
\affil[30]{Institute of Biotechnology, RWTH Aachen University, Worringerweg 3, 52074 Aachen, Germany}
\affil[31]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
\affil[32]{VTT Technical Research Centre of Finland, Espoo, Finland}

@@

%\pagebreak
\section*{Processing results in the application layer}

@@

\pagebreak
\section*{Correlations of area per lipid and thickness with order parameters and form factors}

@@

\pagebreak
\section*{Dependence of form factor and order parameters on the simulation box size}

@@

\pagebreak
\section*{Finding the best models for PC and PE mixtures}

@@

\pagebreak
\section*{Testing machine learning models against literature data}
\begin{figure}[hb]
    \centering
    \includegraphics[width=45mm]{Figures/MLtest/LinearRegressiontest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/Ridgetest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/MLPRegressortest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/Lassotest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/ElasticNettest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/DecisionTreeRegressortest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/RandomForestRegressortest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/GradientBoostingRegressortest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/AdaBoostRegressortest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/KNeighborsRegressortest.pdf}
    \includegraphics[width=45mm]{Figures/MLtest/XGBRegressortest.pdf}
    \caption{Predictions of area per lipids of multi-component membranes composed of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from linear regression model against literature data from simulations using CHARMM36 parameters (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}). Mean squared error (RMSE) and Pearson correlation coefficient are shown for each model. The black line depicts x=y where the prediction equals the literature value. Our machine learning models that are trained using simulations with wide range of force field parameters  have general tendency to predict larger area per lipid values than reported from CHARMM36 simulations in the literature. One explanation could be lower area per lipids predicted by the CHARMM36 parameters when compared with many other force fields in the training set~\cite{javanainen23}, yet other explanations cannot be ruled out.}
    \label{fig:MLTest}
\end{figure}

\pagebreak
\section*{Water permeation through membranes}

@@

\pagebreak
\section*{Water diffusion along membranes}

@@

%\newpage
\pagebreak
\section*{Databank content}

@@

\clearpage
%\section{Experimental data}
\section*{NMR experiments}
Acyl chain order parameters of POPE (Figs.~\ref{POPEexp1} and~\ref{POPEexp2}) and POPG (Figs.~\ref{POPGexp1} and~\ref{POPGexp2}) were analyzed from the same data that were previously recorded to determine headgroup order parameters~\cite{bacle21}. The analysis of the crowded spectral region at 29--31 ppm was based on the previous assignment reported for POPC membranes~\cite{ferreira13}. To measure the order parameters for DOPC (Fig.~\ref{fig:DOPCexp}), the sample was prepared and experiments performed similarly to previous studies~\cite{bacle21}.
